---
title: "Total biomass calculations"
author: "M Siple"
date: "last updated 5/12/22"
output: 
  bookdown::html_document2:
    theme: cerulean
    highlight: tango
bibliography: design-based-indices.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr) #for tables
library(kableExtra) #for tables
options(knitr.table.format = "latex")
```

# Stratum level calculations
Variables for all the following equations are defined in Table \@ref(tab:variable-names-tab)). Catch per unit effort (CPUE) for each tow is calculated as the total catch (in kg) divided by the effort (area swept in $km^2$). The quick calculation for this from the RACEBASE tables is `effort = distance_fished * (0.001 * net_width)`. 

## Stratum CPUE
Represented in Oracle as the `CPUE` table in the GOA or AI schema. Calculated here in the 02_get_cpue.R script.

In this project, CPUE for each haul is calculated in the `01_cleanup_data.R` script:

```{r eval=FALSE}
x <- dat %>%
    mutate(
      wgtcpue = catch_kg / AreaSwept_km2,
      numcpue = number_fish / AreaSwept_km2
    )
```
Where `x` and `dat` are each a table where each row is one haul. 

Represented in Oracle in the `BIOMASS_STRATUM` table in the GOA or AI schema.The weighted mean CPUE for each stratum is:
$$
\overline{CPUE}{_{ik}} =  \frac{ \sum_{j=1}^{n_i} CPUE_{ijk}}{n_i}
$$

Where $i$ is the stratum, $k$ is the species, $j$ is the station, and $n$ is the number of stations. CPUE can be in either numbers or in weight ($kg \cdot km^-2$). 

The variance of the mean CPUE is:

$$
VAR \overline{CPUE}{_{ik}} =  \frac{ \sum_{j=1}^{n_i} (CPUE_{ijk}-\overline{CPUE}_{i,k})^2}{n_i (n_i-1)} 
$$

This is calculated by taking the mean of the weight CPUE and the numeric CPUE:
```{r eval=FALSE}
x %>%
    group_by(year, stratum) %>%
    dplyr::summarize(
      haul_count = length(unique(stationid)), # number of total abundance hauls
      mean_wgt_cpue = mean(wgtcpue),
      var_wgt_cpue = ifelse(haul_count <= 1, NA, var(wgtcpue) / haul_count),
      mean_num_cpue = mean(numcpue),
      var_num_cpue = ifelse(haul_count <= 1, NA, var(numcpue) / haul_count),
      catch_count = length(which(catch_kg > 0)) # number of hauls with nonzero catch
    )
```

## Stratum biomass

Mean biomass in stratum $i$ for species $k$ is estimated as:
$$
\hat{B}_{ik} = \bigg(\frac{A_i}{0.01}\bigg)^2 \cdot \overline{CPUE}{_{ik}}
$$
Where $A_i$ is the area of the $i$th stratum. 

This is represented in the `get_biomass_stratum()` function as follows, where `mean_wgt_cpue` is $\overline{CPUE}{_{ik}}$ and `area` is $A_i$:
```{r eval=FALSE}
x2 <- x %>%
    group_by(year, stratum) %>%
    dplyr::summarize(
      haul_count = length(unique(stationid)), # number of total abundance hauls
      mean_wgt_cpue = mean(wgtcpue),
      var_wgt_cpue = ifelse(haul_count <= 1, NA, var(wgtcpue) / haul_count),
      mean_num_cpue = mean(numcpue),
      var_num_cpue = ifelse(haul_count <= 1, NA, var(numcpue) / haul_count),
      catch_count = length(which(catch_kg > 0)) # number of hauls with nonzero catch
    ) %>%
    dplyr::ungroup() %>%
    select(
      year, stratum,
      haul_count, catch_count,
      mean_wgt_cpue, var_wgt_cpue,
      mean_num_cpue, var_num_cpue
    )
  
  if (all(x2$catch_count <= x2$haul_count)) {
    print("Number of hauls with positive catches is realistic.")
  }
  
  biomass_stratum <- x2 %>%
    dplyr::left_join(strata) %>%
    rowwise() %>% # for applying ifelse() by row
    mutate(
      stratum_biomass = area * mean_wgt_cpue / vulnerability * 0.001, # kg --> mt
      stratum_ratio = area / At,
      biomass_var = area^2 * var_wgt_cpue * 1e-6, # kg--> mt, square it because it's variance
      qt_size = ifelse(haul_count <= 1, 0, 
                       qt(p = 0.025, df = haul_count - 1, lower.tail = F)),
      min_biomass = stratum_biomass - qt_size * sqrt(biomass_var),
      max_biomass = stratum_biomass + qt_size * sqrt(biomass_var),
      stratum_pop = area * mean_num_cpue,
      pop_var = area^2 * var_num_cpue,
      min_pop = stratum_pop - qt_size * sqrt(pop_var),
      max_pop = stratum_pop + qt_size * sqrt(pop_var)
    ) %>%
    mutate(
      min_biomass = ifelse(min_biomass < 0, 0, min_biomass),
      min_pop = ifelse(min_pop < 0, 0, min_pop)
    ) %>% # set low CI to zero if it's negative
    select(survey, year, stratum, stratum_ratio, haul_count, catch_count, mean_wgt_cpue, var_wgt_cpue, mean_num_cpue, var_num_cpue, stratum_biomass, biomass_var, min_biomass, max_biomass, stratum_pop, pop_var, min_pop, max_pop, area) %>%
    mutate(
      Ni = area / 0.01,
      fi = (Ni * (Ni - haul_count)) / haul_count
    )

```

Where `x` is the table of CPUE by stratum that results from `get_cpue()`. 


# Total area calculations

## Total survey CPUE
The total CPUE is the weighted sum of CPUE across all strata, weighted by the total area of each stratum (equation 4 in Wakabayashi et al. 1985). The overall mean CPUE for the entire survey area is $\overline{CPUE}_{Tk}$:

$$
\overline{CPUE}_{Tk} = \frac{\sum_{i}{\overline{CPUE}_{ik}}\cdot A_i}{A_T}
$$

Where $k$ is the species, $i$ is the stratum, $A_i$ is the stratum area, and $A_T$ is the total survey area.

## Total survey biomass
Total area biomass is calculated by summing the strata biomasses:
$$
\hat{B}_{Tk} = \sum_{i}{\hat{B}_{ik} }
$$
Variance for the total area biomass is calculated by summing the variances across strata:
$$
VAR \space \hat{B}_{Tk} = \sum_{i}{VAR \space \hat{B}_{ik} }
$$
The confidence intervals for total area biomass are calculated as:
$$
\hat{B}_{Tk} \mp \space t(\alpha,n_e)\sqrt{\sum_{i} VAR \space \hat{B}_{ik}}
$$
And the degrees of freedom $n_e$ is:

$$
n_e = \frac{(\sum_i f_i \cdot VAR \space CPUE_{i,j,k})^2}{    \sum_i {\frac{f_i^2\cdot ( VAR \space CPUE_{i,j,k})^2}{n_i -1}}}
$$
As calculated in Cochran (1962). This equation is corrected from Equation 12 in Wakabayashi et al. [-@wakabayashi_methods_1985].

The number of individuals of a given species in a given stratum is estimated as:
$$
\hat{P_{i,k}} = \bigg(\frac{A_i}{0.01}\bigg) \cdot \frac{\sum_{j=1}^{n_i} \hat{S}_{i,j,k}} {n_{i,j,k}}
$$
where $\hat{S}_{i,j,k} $ is $\frac{CPUE_{i,j,k}}{w_{i,j,k}}$, an estimate of the number of individuals caught per hectare at station *j* (and $w_{i,j,k}=r_{i,j,k}/s_{i,j,k}$, an estimate of the mean weight of individuals in the subsample). 

## Total survey numbers
The total number of individuals in the survey area is the sum of the stratum population numbers:
$$
\hat{P_{T,k}} = \sum_i \hat{P_{i,k}}
$$

## Size composition
Length frequencies for each station are estimated by expanding the length-frequency subsample to the total catch (per sampling unit):

$$
\hat{S}_{i,j,k,l,m} = {s}_{i,j,k,l,m} \cdot \frac{\hat{S}_{i,j,k}}{\sum_{j=1}^{3} \sum_{l=1}^{L_i,j,k,m} \hat{S}_{i,j,k,l,m}}
$$

where $s_{i,j,k,l,m}$ is the number of individuals within the length-frequency subsample of species $k$.

## Age composition
The number of individuals within each sex and size class for each stratum is:

$$
\hat{P}_{i,j,k,l,m} = \hat{P}_{i,k} \cdot \frac{\sum_{j=1}^{n_i}\hat{S}_{i,j,k,l,m}}{\sum_{j=1}^{n_j} \sum_{m=1}^{L_i,j,k,m} \hat{S}_{i,j,k,l,m}}
$$
where $n_i$ is the number of stations at which length-frequency samples were obtained for species $k$. By summing across sex and/or stratum, from $\hat{P}_{i,j,k,l,m}$, we calculate:
a) Size composition by sex and combined sexes of species $k$ by stratum and for all strata
b) The numbers of species $k$ by sex category for each stratum and combined strata.


# Tables
```{r variable-names-tab, echo=FALSE}
x <- data.frame(symbol = c("A", "B", "c", "CPUE", "D", "F","L","P","S","VAR", "W","a","b","n","$n_e$","p","r","s","t","w","i","j","k","l","m","g","T"),
                definition = c("Geographical area",
                               "Population weight",
                               "Vulnerability",
                               "Catch per unit effort (catch per area trawled)",
                               "Computed distance trawled",
                               "Relative fishing power correction factor",
                               "Number of length categories",
                               "Population size",
                               "Number of individuals within a standard sampling unit (haul)",
                               "Variance", #10
                               "Observed catch weight",
                               "Coefficient of length-weight relationship",
                               "Exponent of the length-weight relationship",
                               "Number of stations",
                               "Effective degrees of freedom",
                               "Trawl path width",
                               "Weight of all individuals in a subsample",
                               "Number of all individuals in a subsample",
                               "Age",
                               "Individual weight", #20
                               "Statistical strata",
                               "Station",
                               "Species",
                               "Length class",
                               "Sex class",
                               "Trawl net",
                               "Total survey area (all strata combined)"), #27
                units = c("$km^2$",
                          "kg",
                          "scalar",
                          "kg/ha",
                          "km",
                          "scalar",
                          "integer",
                          "integer",
                          "number/ha",
                          "unitless",
                          "kg",
                          "$g/cm^-3$",
                          "unitless",
                          "integer",
                          "unitless",
                          "km","kg","integer","years","kg",rep(NA,7)))
kableExtra::kbl(x,
                escape = FALSE,
                caption = "Variable names used in these calculations") %>%
  kableExtra::pack_rows("Subscripts",21,27) %>%
  kableExtra::column_spec(c(1,3),italic = T) %>%
  kable_styling()
```

# References