---
title: "Example: Production Run with Species Complexes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example: Production Run with Species Complexes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here is an EBS example of how you would calculate design-based indices for a species complex, i.e., an aggregate of multiple SPECIES_CODE values. This script produces tables equivalent to HAEHNR.CPUE_EBS_PLUSNW_GROUPED, HAEHNR.BIOMASS_EBS_PLUSNW_GROUPED, HAEHNR.SIZECOMP_EBS_PLUSNW_GROUPED.

```{r setup, eval = F}
library(gapindex)

## Connect to Oracle. Make sure you are on the NOAA internal network or VPN
channel <- gapindex::get_connected()

## Skate SPECIES_CODE values excluding SPECIES_CODE values for egg cases
skate_unid <- RODBC::sqlQuery(
  channel = channel, 
  query = "SELECT SPECIES_CODE
  FROM GAP_PRODUCTS.TEST_SPECIES_CLASSIFICATION 
  WHERE SURVEY_SPECIES = 1 
  AND SPECIES_CODE BETWEEN 400 AND 495 
  AND SPECIES_CODE NOT IN (402, 403, 411, 421, 436, 441, 446, 456,
  461, 473, 474, 476, 478, 481, 484, 486)")

## Pull data. Note the format of the `spp_codes` argument with the GROUP column
production_data <- get_data(
  year_set = 1982:2023,
  survey_set = "EBS",
  spp_codes = rbind(
    ## Skate unident.
    data.frame(GROUP = 400, SPECIES_CODE = skate_unid),
    ## ATF + Kams
    data.frame(GROUP = 10111, SPECIES_CODE = 10110:10112),
    ## FHS + Bering flounder
    data.frame(GROUP = 10129, SPECIES_CODE = 10130:10140),
    ## rock sole unid.
    data.frame(GROUP = 10260, SPECIES_CODE = 10260:10262),
    ## Octopididae
    data.frame(GROUP = 78010, SPECIES_CODE = 78010:78455),
    ## Squid unid
    data.frame(GROUP = 79000, SPECIES_CODE = 79000:79513),
    ## Some test single species
    data.frame(GROUP = c(21720, 30060), SPECIES_CODE = c(21720, 30060))
  ),
  pull_lengths = TRUE, 
  haul_type = 3, 
  abundance_haul = "Y",
  channel = channel)

## Zero-fill and calculate CPUE
production_cpue <- calc_cpue(racebase_tables = production_data)

## Calculate Biomass, abundance, mean CPUE, and associated variances by stratum
production_biomass_stratum <- 
  gapindex::calc_biomass_stratum(racebase_tables = production_data,
                                 cpue = production_cpue)

## Aggregate Biomass to subareas and region
production_biomass_subarea <- 
  calc_biomass_subarea(racebase_tables = production_data, 
                       biomass_strata = production_biomass_stratum)

## Calculate size composition by stratum. Note fill_NA_method == "BS" because
## our region is EBS, NBS, or BSS. If the survey region of interest is AI or 
## GOA, use "AIGOA". See ?gapindex::gapindex::calc_sizecomp_stratum for more
## details. 
production_sizecomp_stratum <- 
  gapindex::calc_sizecomp_stratum(
    racebase_tables = production_data,
    racebase_cpue = production_cpue,
    racebase_stratum_popn = production_biomass_stratum,
    spatial_level = "stratum",
    fill_NA_method = "BS")

## Aggregate size composition to subareas/region
production_sizecomp_subarea <- gapindex::calc_sizecomp_subarea(
  racebase_tables = production_data,
  size_comps = production_sizecomp_stratum)


## rbind stratum and subarea/region biomass estimates into one dataframe
names(x = production_biomass_stratum)[
  names(x = production_biomass_stratum) == "STRATUM"
] <- "AREA_ID"
production_biomass <- rbind(production_biomass_stratum, 
                            production_biomass_subarea)

## rbind stratum and subarea/region biomass estimates into one dataframe
names(x = production_sizecomp_stratum)[
  names(x = production_sizecomp_stratum) == "STRATUM"] <- "AREA_ID"
production_sizecomp <- 
  rbind(production_sizecomp_subarea,
        production_sizecomp_stratum[, names(production_sizecomp_subarea)])
```

## Equivalent Tables:

HAEHNR.CPUE_EBS_PLUSNW_GROUPED: `production_cpue`

HAEHNR.BIOMASS_EBS_PLUSNW_GROUPED: `production_biomass`

HAEHNR.SIZECOMP_EBS_PLUSNW_GROUPED: `production_sizecomp`
